ray

### 2022-03-19 outputs/paper/ray.Rmd
more graphs

>>>>

EDA

```{r}

# % missing hlthmntl entries 0.09
n <- length(raw_data$hlthmntl)
length(which(is.na(raw_data$hlthmntl)))/n

```{r}
class_data <- raw_data %>% select(c(hlthmntl, class)) %>% filter(!is.na(class), !is.na(hlthmntl)) %>%
  mutate(class_d = case_when(
    class == 1 ~ "lower",
    class == 2 ~ "working",
    class == 3 ~ "lower middle",
    class == 4 ~ "middle",
    class == 5 ~ "upper middle",
    class == 6 ~ "upper",

  )) %>%
  mutate(mntl_poor = case_when(
    hlthmntl <= 3 ~ 'Good',
    hlthmntl >= 4 ~ 'Not Good',
  ))

class_data %>% ggplot(aes(x=hlthmntl, fill=class_d)) +
  geom_histogram(aes(y = ..density..), position = 'dodge')

class_data %>% ggplot(aes(x= mntl_poor,  group=class_d))+
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
  geom_text(aes( label = scales::percent(..prop..),
                 y= ..prop.. ), stat= "count", vjust = -.5) +
  labs(y = "Percent", fill="mntl_poor") +
  facet_grid(~class_d) +
  scale_y_continuous(labels = scales::percent)
```

```{r}
income_data <- raw_data %>% select(c(hlthmntl, income16)) %>% filter(!is.na(income16), !is.na(hlthmntl)) %>%
  mutate(income = case_when(
    income16 <= 8 ~ "lowest",
    income16 <= 12 ~ "lower",
    income16 <= 17 ~ "lower middle",
    income16 <= 19 ~ "middle",
    income16 <= 21 ~ "upper middle",
    income16 > 21 ~ "upper",

  )) %>%
  mutate(mntl_poor = case_when(
    hlthmntl <= 3 ~ 'Good',
    hlthmntl >= 4 ~ 'Not Good',
  ))

income_data %>% ggplot(aes(x=hlthmntl, fill=income)) +
  geom_histogram(aes(y = ..density..), position = 'identity', alpha=0.3)

income_data %>% ggplot(aes(x= mntl_poor,  group=income))+
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
  geom_text(aes( label = scales::percent(..prop..),
                 y= ..prop.. ), stat= "count", vjust = -.5) +
  labs(y = "Percent", fill="mntl_poor") +
  facet_grid(~income) +
  scale_y_continuous(labels = scales::percent)
```

```{r}
first_born <- raw_data %>% select(c(hlthmntl, agekdbrn)) %>% filter(!is.na(agekdbrn), !is.na(hlthmntl)) %>%
  mutate(born = case_when(
    agekdbrn <= 18 ~ "1",
    agekdbrn <= 25 ~ "2",
    agekdbrn <= 30 ~ "3",
    agekdbrn <= 40 ~ "4",
    agekdbrn <= 50 ~ "5",
    agekdbrn > 50 ~ "6",

  )) %>%
  mutate(mntl_poor = case_when(
    hlthmntl <= 3 ~ 'Good',
    hlthmntl >= 4 ~ 'Not Good',
  ))

first_born %>% ggplot(aes(x=hlthmntl, fill=born)) +
  geom_histogram(aes(y = ..density..), position = 'dodge')

first_born %>% ggplot(aes(x= mntl_poor,  group=born))+
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
  geom_text(aes( label = scales::percent(..prop..),
                 y= ..prop.. ), stat= "count", vjust = -.5) +
  labs(y = "Percent", fill="mntl_poor") +
  facet_grid(~born) +
  scale_y_continuous(labels = scales::percent)
```
